Local Point Spread Function Approximation of High-Rank Hessians in Inverse Problems Governed by PDEs\thanks{Research supported partially by NSF-DMS-CAREER-1654311 and NSF-OAC-1550547.}

### 2024-03-05 presentations/siampp24-tucker/talk.tex
many updates to SIAMPP24 slides

\begin{frame}
	\frametitle{Outline}
	TODO: edit!!!

	\begin{itemize}
		\item Motivation
		\vspace{0.02in}
		\begin{itemize}
			\item Why we need predictive models with quantified uncertainties
			to accurately anticipate future sea level rise?
			\vspace{0.02in}
			\item Why we need more efficient approximations of high rank
			Hessians?
		\end{itemize}
		\vspace{0.02in}
		\item Bayesian inverse problems governed by PDEs (preliminaries)
		\vspace{0.02in}
		\item Exploiting low-dimensional structure in Bayesian inverse problems governed by PDEs
		\vspace{0.02in}
		\begin{itemize}
			\item (Global) low rank approximation
			\vspace{0.02in}
			\item Hierarchical off-diagonal low rank (HODLR) approximation
			\vspace{0.02in}
			\item Point spread function approximation combined with
			hierarchical matrix (H-matrix) approximation
		\end{itemize}
		\vspace{0.02in}
		\item Example: Inversion for the basal friction
		coefficient in an ice sheet flow problem
		\vspace{0.02in}
		\item Conclusions
	\end{itemize}

@@

				\vspace{0.1in}
				where $\bs{V}_r$ and $\bs{\Lambda}_r$ are the eigenvectors/values of
				%  the eigenvalue problem 
				$\Hmisfit \bs{v}_i = \lambda_i \prcov^{-1} \bs{v}_i $
				%, \bs{\Lambda}_r = \diag(\lambda_i)_{i=1}^r$
				%  are truncated
				%  generalized eigenvectors of the data misfit
				%  Hessian, $\bs{D}_r =\diag(\lambda_i/(\lambda_i+1))$, and used
				%  the Sherman-Morrison-Woodbury th. % to invert/factor.perwidth}
			\end{itemize} 
		\end{column}
		\begin{column}{0.49\paperwidth}
			\begin{itemize}
				\item<2-> Spectrum of the prior-preconditioned data-misfit Hessian for Antarctica

@@

}
			\vspace{-0.1in}
		\end{column}
		\begin{column}{0.25\columnwidth}
			\only<2->
			{
				\begin{figure} 	
					\begin{tikzpicture}
					\begin{scope}[yscale=0.2, xscale=0.2]
					\pgfmathtruncatemacro{\L}{3}
					\pgfmathtruncatemacro{\scalefac}{16/(2^\L)}
					\pgfmathtruncatemacro{\shiftval}{2^(\L-1)}
					\begin{scope}[yscale=-1, xscale=1]
					\begin{scope}[yshift=-\shiftval, xshift=-\shiftval]
					\begin{scope}[yscale=\scalefac, xscale=\scalefac]
					\pgfmathtruncatemacro{\msize}{2^\L}
					%\pgfmathtruncatemacro{\gridstep}{2^(-\L)}
					\pgfmathtruncatemacro{\colstep}{40/(\L-1)}
					%\draw[step=\gridstep cm, gray, very thin] (0, 0) grid (\msize, \msize);
					\foreach \l in {1,...,\L}
					{
						\pgfmathtruncatemacro{\delI}{\msize*2^(-\l)}
						\pgfmathtruncatemacro{\DelI}{2*\delI}
						%\pgfmathtruncatemacro{\colorl}{50+(\l-1)*\colstep};
						\pgfmathtruncatemacro{\colorl}{75}
						\pgfmathtruncatemacro{\maxi}{2^(\l-1)}
						\foreach \i in {1,...,\maxi}
						{
							\pgfmathtruncatemacro{\a}{(\i-1)*\DelI}
							\pgfmathtruncatemacro{\b}{(\i-1)*\DelI+\delI}
							\pgfmathtruncatemacro{\c}{\i*\DelI}
							\ifnum\l=1
							\filldraw[fill=teal] (\a, \b) rectangle (\b, \c);
							\filldraw[fill=teal] (\b, \a) rectangle (\c, \b);
							\else
							\ifnum\l=2 
							\filldraw[fill=olive] (\a, \b) rectangle (\b, \c);
							\filldraw[fill=olive] (\b, \a) rectangle (\c, \b);
							\else
							\filldraw[fill=gray] (\a, \b) rectangle (\b, \c);
							\filldraw[fill=gray] (\b, \a) rectangle (\c, \b);
							\fi
							\fi 
						}
					}
					% diagonal blocks
					\pgfmathtruncatemacro{\delI}{\msize*2^(-\L)}
					\pgfmathtruncatemacro{\DelI}{2*\delI}
					\pgfmathtruncatemacro{\colorl}{20+(\L-1)*\colstep}
					\pgfmathtruncatemacro{\maxi}{2^(\L-1)}
					\foreach \i in {1,...,\maxi}
					{
						\pgfmathtruncatemacro{\a}{(\i-1)*\DelI}
						\pgfmathtruncatemacro{\b}{(\i-1)*\DelI+\delI}
						\pgfmathtruncatemacro{\c}{\i*\DelI}
						\filldraw[fill=violet] (\a, \a) rectangle (\b, \b);
						\filldraw[fill=violet] (\b, \b) rectangle (\c, \c);
					}
					\end{scope}
					\end{scope}
					\end{scope}
					\end{scope}
					\end{tikzpicture}
					\caption{HODLR matrix partition structure}
				\end{figure} 
			}

@@

%% -------------------------------------------------------------------
\begin{frame}
	\frametitle{Hierarchical ($\mathcal{H}$-) matrices}
	\framesubtitle{Construction from matrix-free operators}
	\begin{itemize}
		\setlength\itemsep{1.5em}
		\item<1-> Classic methods for building $\mathcal{H}$-matrices \only<9>{\textbf{\underline{require the evaluation of matrix entries}}}\only<1-8>{require the evaluation of matrix entries} $\bs{H}_{i,j}$
		\item<2-> Purely algebraic methods based on \textit{peeling} can generate $\mathcal{H}$-matrices of matrix-free operators via the application of specially chosen (e.g., random) vectors

		{\scriptsize{P.G. Martinsson.
			{\em Compressing rank-structured matrices via randomized sampling},
			SIAM Journal on Scientific Computing, (2016).}
		}
		%\begin{itemize}
		%	\item {\scriptsize T. Hartland, G. Stadler, M. Perego, K. Li\'{e}geois and N. Petra. \emph{Hierarchical off-diagonal approximation of Hessians in inverse problems}. To be submitted.}
		%	\item Tucker Hartland is giving a talk about this later
		%	\item 3:00-3:25 Thursday, MS140, Room: Augusta F - 7th Floor
		%\end{itemize} 
		\item<3-> \alert{Challenge:} peeling process can be better than low rank, but \textbf{still expensive} \onslide<4->{and expense grows with the complexity of hierarchical matrix partitioning}
		\item<5-> \alert{Approach:} speed up the $\mathcal{H}$-matrix construction by
		exploiting the following problem structure
		\begin{itemize}
			\vspace{0.1in}
			\setlength\itemsep{0.5em}
			\item<6-> \textbf{Localized sensitivities}
			\item<7-> \textbf{Scaled local mean-displacement invariance}
			\item<8-> \textbf{Non-negative impulse responses}
		\end{itemize}
	\end{itemize}
\end{frame}

%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Local mean-displacement invariance}
	\begin{center}
		\textbf{Impulse response $\phi_x$:}
		\begin{align*}
		\phi_x := H_d~ \delta_x =& \text{ action of Hessian operator on delta distribution at x} \\
		\mu(x) :=& \text{ center of mass (mean) of }\phi_x
		\end{align*}

@@

\begin{frame}

@@

%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Hessian approximation method: big idea}
	\begin{itemize}
		\item {\bf Step 1:} Compute ``batches'' of impulse responses by applying Hessian to Dirac combs
		\item {\bf Step 2:} Interpolate known impulse responses to approximate unknown Hessian entries $\mathbf{H}_{ij}$
		\item {\bf Step 3:} Convert to $\mathcal{H}$-matrix to do linear algebra
	\end{itemize}
	\includegraphics[width=0.45\columnwidth]{impulse_batch1.png}  \includegraphics[width=0.45\columnwidth]{impulse_batch2.png}
	*Impulse response batches from Ice Mountain
\end{frame}
%---------------------------------------------------------------------------------%
%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Technical details}
	{\Large
		\begin{itemize}
			\setlength\itemsep{2em}
			\item How do we choose the impulse response points?
			\begin{itemize}
				\item {\large How do we make sure they don't overlap?}
			\end{itemize}
			\item How do we interpolate the impulse responses? 
			\begin{itemize}
				\item {\large What about boundary issues?}
			\end{itemize}
		\end{itemize}
	}
\end{frame}
%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{How to choose impulse response points?}
	One hessian matrix-vector product $\rightarrow$ many impulse responses
	\begin{center}
		\includegraphics[width=0.5\columnwidth]{impulse_batch2.png} 
	\end{center}
	\begin{itemize}
		\item \textbf{Goal:} choose as many points as possible, such that the impulse response supports don't overlap
		\item \textbf{Dilemma:} How can we know the impulse response supports before we compute them?
	\end{itemize}
\end{frame}
%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Matrix analogy: getting all row sums}

@@

	$$\mathbf{A}^T \begin{bmatrix}
	1 \\ 1 \\ \vdots \\ 1
	\end{bmatrix} = \begin{bmatrix}
	\text{sum of }\mathbf{A} \text{ col }1 \\
	\text{sum of }\mathbf{A} \text{ col }2 \\
	\vdots \\
	\text{sum of }\mathbf{A} \text{ col }N
	\end{bmatrix}$$

	Apply matrix to vector of ones $\rightarrow$ get row sums for all rows

@@

	$$(H_d^T C)^*(y) = \int_\Omega \left(H_d \delta_y\right)(x) dx$$

	Apply Hessian to constant function $\rightarrow$ get volumes of every impulse response
\end{frame}
%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Mean and standard deviations of impulse responses}
	\begin{itemize}
		\setlength\itemsep{2em}
		\item Let $V(x)$, $\mu(x)$, and $\Sigma(x)$ be the ``volume'', ``mean'', and ``variance'' of $\phi_x$
		\item Let $C$, $L^i$, and $Q^{ij}$ be the following functions:
		\begin{equation*}
		C(x) := 1, \qquad
		L^i(x) := x^i, \qquad
		Q^{ij}(x) = x^i x^j
		\end{equation*}
		\item Then
		\begin{align*}
		V =& \left(H_d^T C\right)^* \\
		\mu^i =& \left(H_d^T L^i\right)^* / V \\
		\Sigma^{ij} =& \left(H_d^T Q^{ij}\right)^* / V - \mu^i\cdot \mu^j
		\end{align*}
		\item Apply Hessian to constant, linear, and quadratic functions $\rightarrow$ get estimates of support for every impulse response
	\end{itemize}

@@

		\item Pack ellipsoids using \textbf{greedy algorithm}
	\end{itemize}
	\begin{center}
		\includegraphics[width=0.5\columnwidth]{impulse_batch2.png} 
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Radial basis function interpolation}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{interpolation_rbf.pdf}
	\end{center}
	\begin{itemize}
		\item Interpolate impulse responses using polyharmonic spline radial basis functions.
		\item Use only $k$-nearest neighbors (must solve $k \times k$ linear system)
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Computational cost}
	\begin{itemize}
		\setlength\itemsep{2em}
		\item \textbf{Hessian-vector products:} 
		$$6 + n_\text{batches}$$
		E.g., $11$ Hessian-vector products for $5$ batches of impulse responses
		\item \textbf{Ellipsoid intersection tests:}
		$$O(Nm)$$
		where $m$ is total number of impulse responses in all batches
		\item \textbf{Elementary operations} to build and use the H-matrix:
		$$O(N \log N)$$ 
	\end{itemize}
\end{frame}

@@

\end{frame}
%---------------------------------------------------------------------
\begin{frame}
	\frametitle{Example: Ice sheet inverse problem}

	\begin{figure}
		%[width=0.8\columnwidth]
		\centering\includegraphics[scale=0.6]{extraplots/reconstructions-ice.png}
		%\centering\includegraphics[scale=0.18]{figures/mstar_0.25noise_cropped.png}
		%\includegraphics[scale=0.18]{figures/mstar_0.05noise_cropped.png}
		%\includegraphics[scale=0.18]{figures/mstar_0.01noise_cropped.png}
	\end{figure}
	\begin{center}
		The log basal sliding parameter computed from the PDE-constrained
		optimization problem with noise levels: $25\%$ (left), $5.0\%$
		(middle), and $1.0\%$ (right).
	\end{center}
\end{frame}
%---------------------------------------------------------------------
\begin{frame}
	\frametitle{Example: Ice sheet inverse problem}
	\framesubtitle{Convergence history for solving the Stokes inverse problem using inexact Newton PCG}

@@

	\footnotesize
	\begin{itemize}
		\item \#CG:  the number of PCG iterations used to solve the Newton system.
		\item \#Stokes:the total number of Stokes PDE solves performed in each Newton iteration.
		\item $\|\mathbf{g}\|$: the $l^2$ norm of the gradient.
	\end{itemize}
\end{frame}
%---------------------------------------------------------------------
\begin{frame}
	\frametitle{Example: ``Ice mountain'' inverse problem}
	\framesubtitle{Convergence history and eigenvalue clustering}

@@

	\begin{itemize}
		\setlength\itemsep{1.5em}
		\item Hessian approximations or preconditioners are essential for solution of Bayesian inverse problems governed by
		partial differential equations.
		\vspace{0.05in}
		\item Low-rank approximations of the Hessian become
		prohibitive as the data becomes more informative, as is the case
		for ice sheet inverse problems.
		\vspace{0.05in}
		\item Local point spread function interpolation combined with Hierarchical matrix representations promise a more efficient
		Hessian approximation.
	\end{itemize}
	\vspace{2em}
	{\small Alger, N., Hartland, T., Petra, N., Ghattas, O. (2024). Point spread function approximation of high rank Hessians with locally supported non-negative integral kernels. To appear in SISC.}
\end{frame}

>>>>

\begin{frame}[t]
	\frametitle{Outline}

	\begin{itemize}
		\setlength\itemsep{1.0em}
		\item Motivation
		\vspace{0.02in}
		\begin{itemize}
			\setlength\itemsep{1.0em}
			\item Predictive continental scale ice sheet models for anticipating future sea level rise with quantified uncertainties
			\vspace{0.02in}
			\item The need to go beyond ``low rank'' Hessian approximations
		\end{itemize}
		%\vspace{0.02in}
		%\item Bayesian inverse problems governed by PDEs (preliminaries)
		\vspace{0.02in}
		\item Exploiting structured low-rank properties of (Bayesian) inverse problems governed by PDEs
		\vspace{0.02in}
		\begin{itemize}
			\setlength\itemsep{1.0em}
			\item (Global) low rank approximation
			\vspace{0.02in}
			\item Hierarchical off-diagonal low rank (HODLR) approximation
			\vspace{0.02in}
			\item A local point spread function method as a means for efficient
			$\mathcal{H}$ierarchical-matrix approximation
		\end{itemize}
		\vspace{0.02in}
		\item Example: Inversion for the basal friction
		coefficient field in an ice sheet flow problem
		\vspace{0.02in}
		\item Conclusions
	\end{itemize}

@@

				\vspace{0.1in}
				where $\bs{V}_r$ and $\bs{\Lambda}_r$ are the eigenvectors/values of
				%  the eigenvalue problem 
				$\prcov^{\top/2}\Hmisfit\prcov^{1/2}$
				%, \bs{\Lambda}_r = \diag(\lambda_i)_{i=1}^r$
				%  are truncated
				%  generalized eigenvectors of the data misfit
				%  Hessian, $\bs{D}_r =\diag(\lambda_i/(\lambda_i+1))$, and used
				%  the Sherman-Morrison-Woodbury th. % to invert/factor.perwidth}
			\end{itemize} 
		\end{column}
		\begin{column}{0.49\paperwidth}
			\begin{itemize}
				\item<2-> Spectrum of the prior-preconditioned data-misfit Hessian for Antarctica

@@

}
			\vspace{-0.1in}
		\end{column}
		\begin{column}{0.25\columnwidth}
			\only<2->
			{
				\begin{figure} 	
					\begin{tikzpicture}
					\begin{scope}[yscale=0.2, xscale=0.2]
					\pgfmathtruncatemacro{\L}{3}
					\pgfmathtruncatemacro{\scalefac}{16/(2^\L)}
					\pgfmathtruncatemacro{\shiftval}{2^(\L-1)}
					\begin{scope}[yscale=-1, xscale=1]
					\begin{scope}[yshift=-\shiftval, xshift=-\shiftval]
					\begin{scope}[yscale=\scalefac, xscale=\scalefac]
					\pgfmathtruncatemacro{\msize}{2^\L}
					%\pgfmathtruncatemacro{\gridstep}{2^(-\L)}
					\pgfmathtruncatemacro{\colstep}{40/(\L-1)}
					%\draw[step=\gridstep cm, gray, very thin] (0, 0) grid (\msize, \msize);
					\foreach \l in {1,...,\L}
					{
						\pgfmathtruncatemacro{\delI}{\msize*2^(-\l)}
						\pgfmathtruncatemacro{\DelI}{2*\delI}
						%\pgfmathtruncatemacro{\colorl}{50+(\l-1)*\colstep};
						\pgfmathtruncatemacro{\colorl}{75}
						\pgfmathtruncatemacro{\maxi}{2^(\l-1)}
						\foreach \i in {1,...,\maxi}
						{
							\pgfmathtruncatemacro{\a}{(\i-1)*\DelI}
							\pgfmathtruncatemacro{\b}{(\i-1)*\DelI+\delI}
							\pgfmathtruncatemacro{\c}{\i*\DelI}
							\ifnum\l=1
							\filldraw[fill=teal] (\a, \b) rectangle (\b, \c);
							\filldraw[fill=teal] (\b, \a) rectangle (\c, \b);
							\else
							\ifnum\l=2 
							\filldraw[fill=olive] (\a, \b) rectangle (\b, \c);
							\filldraw[fill=olive] (\b, \a) rectangle (\c, \b);
							\else
							\filldraw[fill=gray] (\a, \b) rectangle (\b, \c);
							\filldraw[fill=gray] (\b, \a) rectangle (\c, \b);
							\fi
							\fi 
						}
					}
					% diagonal blocks
					\pgfmathtruncatemacro{\delI}{\msize*2^(-\L)}
					\pgfmathtruncatemacro{\DelI}{2*\delI}
					\pgfmathtruncatemacro{\colorl}{20+(\L-1)*\colstep}
					\pgfmathtruncatemacro{\maxi}{2^(\L-1)}
					\foreach \i in {1,...,\maxi}
					{
						\pgfmathtruncatemacro{\a}{(\i-1)*\DelI}
						\pgfmathtruncatemacro{\b}{(\i-1)*\DelI+\delI}
						\pgfmathtruncatemacro{\c}{\i*\DelI}
						\filldraw[fill=violet] (\a, \a) rectangle (\b, \b);
						\filldraw[fill=violet] (\b, \b) rectangle (\c, \c);
					}
					\end{scope}
					\end{scope}
					\end{scope}
					\end{scope}
					\end{tikzpicture}
					\caption{HODLR matrix structure}
				\end{figure} 
			}

@@

%% -------------------------------------------------------------------
\begin{frame}[t]
	\frametitle{Hierarchical ($\mathcal{H}$-) matrices}
	\framesubtitle{Construction from matrix-free operators}
	\begin{columns}
	\begin{column}{0.59\columnwidth}
	\begin{itemize}
		\setlength\itemsep{1.0em}
		\item<1-> Classic methods for building $\mathcal{H}$-matrices \\ \only<8>{\textbf{\underline{require evaluation of matrix entries}}}\only<1-7>{require evaluation of matrix entries} $\bs{H}_{i,j}$
		\item<2-> Purely algebraic methods based on \textit{peeling} can generate $\mathcal{H}$-matrices of matrix-free operators via the application of specially chosen (e.g., random) vectors
		{\scriptsize{P.G. Martinsson.
			{\em Compressing rank-structured matrices via randomized sampling},
			SIAM Journal on Scientific Computing, (2016).}
		}
	\end{itemize}
		%\begin{itemize}
		%	\item {\scriptsize T. Hartland, G. Stadler, M. Perego, K. Li\'{e}geois and N. Petra. \emph{Hierarchical off-diagonal approximation of Hessians in inverse problems}. To be submitted.}
		%	\item Tucker Hartland is giving a talk about this later
		%	\item 3:00-3:25 Thursday, MS140, Room: Augusta F - 7th Floor
		%\end{itemize} 
\end{column}
\begin{column}{0.4\columnwidth}
\begin{figure} 
\begin{center} 
\includegraphics[scale=0.16]{Hmatrix_Structure.png}
\caption{$\mathcal{H}$-matrix structure}
\end{center}
\end{figure} 
\btVFill
\end{column}
\end{columns}
\begin{itemize}
	%\vspace{0.1in}
		\item<3-> \alert{Challenge:} peeling process can be better than low rank, but \textbf{still expensive} \onslide<4->{and expense grows with the complexity of hierarchical matrix partitioning}
\item<5-> \alert{Approach:} speed up the $\mathcal{H}$-matrix construction by
exploiting additional problem structure: \only<6->{\textbf{localized sensitivities}} \only<7->{and (local mean-)\textbf{displacement invariance}}
\end{itemize}

%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Local mean-displacement invariance}
	\begin{center}
		\textbf{Impulse response $\phi_x$:}
		\begin{align*}
		\phi_x := H_{\text{\tiny misfit}}~ \delta_x =& \text{ action of data-misfit Hessian operator on delta distribution at x} \\
		\mu(x) :=& \text{ center of mass (mean) of }\phi_x
		\end{align*}

@@

\begin{frame}[t]

@@

%---------------------------------------------------------------------------------%
\begin{frame}[t]
	\frametitle{Hessian approximation method: big idea}
	\begin{itemize}
		\item {\bf Step 1:} Compute ``batches'' of impulse responses by applying Hessian to Dirac combs
		\item {\bf Step 2:} Interpolate known impulse responses to approximate unknown Hessian entries $\mathbf{H}_{ij}$
		\item {\bf Step 3:} Convert to $\mathcal{H}$-matrix to do linear algebra (with low computational complexity)
	\end{itemize}
    \begin{figure} 
	\includegraphics[width=0.375\columnwidth]{impulse_batch1.png} \hfil  
	\includegraphics[width=0.375\columnwidth]{impulse_batch2.png}
	\end{figure} 
	*Impulse response batches from Ice Mountain
\end{frame}
%---------------------------------------------------------------------------------%

%---------------------------------------------------------------------------------%
\begin{frame}[t]
	\frametitle{How to choose impulse response points?}

	\begin{columns}
		\begin{column}{0.5\columnwidth}
			One Hessian-vector product can yield many impulse responses (with small region of support)
		\begin{center}
			\includegraphics[width=0.75\columnwidth]{impulse_batch2.png} 
		\end{center}
	    \end{column}
    \begin{column}{0.49\columnwidth}
    	\only<2>
    	{ 
		\begin{itemize}
			\item \textbf{Goal:} determine as many points as possible, such that the associated impulse response supports do not overlap
			\item \textbf{Dilemma:} how can we know the impulse response supports before we compute them?
		\end{itemize}
	    }
	\end{column} 
	\end{columns}

\end{frame}
%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Matrix analogy: getting all row sums}

@@

	$$\mathbf{A}^{\top} \begin{bmatrix}
	1 \\ 1 \\ \vdots \\ 1
	\end{bmatrix} = \begin{bmatrix}
	\text{sum of }\mathbf{A} \text{ col }1 \\
	\text{sum of }\mathbf{A} \text{ col }2 \\
	\vdots \\
	\text{sum of }\mathbf{A} \text{ col }N
	\end{bmatrix}$$

	Sum of each row obtained by transpose apply to vector of ones

@@

	$$(H_{\text{\tiny misfit}}^{\dagger} C)^*(y) = \int_\Omega \left(H_{\text{\tiny misfit}} \delta_y\right)(x) \,\mathrm{d}x=\int_{\Omega}\phi_{y}(x)\,\mathrm{d}x$$

	Volume of each impulse response obtained by adjoint Hessian operator apply to constant function
\end{frame}
%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Mean and standard deviations of impulse responses}
	\begin{itemize}
		\setlength\itemsep{2em}
		\item Let $V(x)$, $\mu(x)$, and $\Sigma(x)$ be the ``volume'', ``mean'', and ``variance'' of $\phi_x$
		\item Let $C$, $L^i$, and $Q^{ij}$ be the following functions:
		\begin{equation*}
		C(x) := 1, \qquad
		L^i(x) := x^i, \qquad
		Q^{ij}(x) = x^i x^j
		\end{equation*}
		\item Then
		\begin{align*}
		V =& \left(H_{\text{\tiny misfit}}^{\dagger} C\right)^* \\
		\mu^i =& \left(H_{\text{\tiny misfit}}^{\dagger} L^i\right)^* / V \\
		\Sigma^{ij} =& \left(H_{\text{\tiny misfit}}^{\dagger} Q^{ij}\right)^* / V - \mu^i\cdot \mu^j
		\end{align*}
		\item Estimates of impulse response supports obtained from 
		first few moments (adjoint Hessian apply to constant, linear, and quadratic functions)
	\end{itemize}

@@

		\item Pack ellipsoids using \textbf{greedy algorithm}
	\end{itemize}
	\begin{center}
		\includegraphics[width=0.375\columnwidth]{impulse_batch2.png} 
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Computational cost}
	\begin{itemize}
		\setlength\itemsep{2em}
		\item \textbf{Hessian-vector products:} 
		$$1+\text{dim}(\Omega)(\text{dim}(\Omega)+3)/2 + n_\text{batches}$$
		E.g., $11$ Hessian-vector products for $5$ batches of impulse responses and $\text{dim}(\Omega)=2$
		\item \textbf{Ellipsoid intersection tests:}
		$$O(Nm)$$
		where $m$ is total number of impulse responses in all batches
		\item \textbf{Elementary operations} to build and use the $\mathcal{H}$-matrix:
		$$O(N \log N)$$ 
	\end{itemize}
\end{frame}

@@

\end{frame}
%---------------------------------------------------------------------
\begin{frame}
	\frametitle{Example: ``Ice mountain'' inverse problem}

	\begin{figure}
		%[width=0.8\columnwidth]
		\centering\includegraphics[scale=0.6]{extraplots/reconstructions-ice.png}
		%\centering\includegraphics[scale=0.18]{figures/mstar_0.25noise_cropped.png}
		%\includegraphics[scale=0.18]{figures/mstar_0.05noise_cropped.png}
		%\includegraphics[scale=0.18]{figures/mstar_0.01noise_cropped.png}
	\end{figure}
	\begin{center}
		The log basal sliding parameter computed from the PDE-constrained
		optimization problem with noise levels: $25\%$ (left), $5.0\%$
		(middle), and $1.0\%$ (right).
	\end{center}
\end{frame}
%---------------------------------------------------------------------
\begin{frame}
	\frametitle{Example: ``Ice mountain'' inverse problem}
	\framesubtitle{Convergence history for solving the Stokes inverse problem using inexact Newton PCG}

@@

	\footnotesize
	\begin{itemize}
		\item \#CG:  the number of PCG iterations used to solve the Newton system.
		\item \#Stokes: the total number of Stokes PDE solves performed in each Newton iteration.
		\item $\|\mathbf{g}\|$: the Euclidean norm of the gradient.
	\end{itemize}
\end{frame}
%---------------------------------------------------------------------
\begin{frame}
	\frametitle{Example: ``Ice mountain'' inverse problem}
	\framesubtitle{Convergence history and eigenvalue clustering}

@@

	\begin{itemize}
		\setlength\itemsep{1.5em}
		\item Hessian approximations or preconditioners are essential for solution of Bayesian inverse problems governed by
		partial differential equations.
		\vspace{0.05in}
		\item Low-rank approximations of the Hessian become
		prohibitive as the data becomes more informative, as is the case
		for continental-scale ice sheet inverse problems.
		\vspace{0.05in}
		\item Local point spread function interpolation combined with hierarchical matrix representations promise a more efficient
		Hessian approximation.
	\end{itemize}
	\vspace{2em}
	{\small Alger, N., Hartland, T., Petra, N., Ghattas, O. (2024). Point spread function approximation of high rank Hessians with locally supported non-negative integral kernels. To appear in SISC.}
\end{frame}
